---
title: "Activity monitor data analysis"
author: "Mingda Ai"
date: "Sunday, January 18, 2015"
output: html_document
---

##Loading and preprocessing the data:

```{r echo = TRUE}
act<-read.csv("activity.csv")
```

##What is mean total number of steps taken per day?:

```{r echo = TRUE}
library(ggplot2)
step_day<-tapply(act$step, act$date, sum)
qplot(x=step_day, xlab="total number of steps")
summary(step_day)[c(3,4)]
```

##What is the average daily activity pattern? 

```{r echo = TRUE}
step_int<-tapply(act$step, act$interval, mean, na.rm = TRUE)
qplot(x=as(names(step_int),"numeric"), y=step_int, geom="line", xlab="5-minute interval", ylab="average number of steps taken")
max_index<-order(step_int, decreasing = TRUE)[1]
```

Interval contains the maximun steps:
```{r echo = TRUE}
names(step_int)[max_index]
```

The total number of missing value is:
```{r echo = TRUE}
sum(is.na(act$step))
```

##Inputing missing values
```{r echo = TRUE}
step_int_data<-data.frame(interval=names(step_int), avg_steps=step_int)
act2<-merge(act, step_int_data, by="interval")
na_lines<-is.na(act2$step)
act2[na_lines,2] = act2[na_lines,4]
step_day2<-tapply(act2$step, act$date, sum)
qplot(x=step_day2, xlab="total number of steps")
summary(step_day2)[c(3,4)]
```
Fill the value with the average steps on the interval across all days.
After inputing the data, the mean value remain the same since we use the average, but median changes.

##Are there differences in activity patterns between weekdays and weekends?
```{r echo = TRUE}
is_week_end <-function(date) {
    week<-weekdays(as.Date(date,"%Y-%m-%d"))
    if(week == "Sunday" || week=="Saturday")
    {
        "weekend"
    }
    else
    {
        "weekday"
    }
}
act2$weekday<-sapply(act2$date, is_week_end)
step_group<-split(act2, act2$weekday)

mean1<-tapply(step_group[[1]]$step, step_group[[1]]$interval, mean, na.rm = TRUE)
mean2<-tapply(step_group[[2]]$step, step_group[[2]]$interval, mean, na.rm = TRUE)
step_mean1<-data.frame(inverval=as.numeric(names(mean1)), mean=mean1, weekday="weekday")
step_mean2<-data.frame(inverval=as.numeric(names(mean2)), mean=mean2, weekday="weekend")
step_mean<-rbind(step_mean1, step_mean2)
p <- ggplot(data = step_mean, mapping = aes(x = inverval, y = mean))
p + geom_line() + facet_wrap(~weekday, ncol=1) + xlab("5-minute interval") + ylab("the average number of steps taken")
```